PANCHANGE_analysis

### 2020-09-03 PANCHANGE_analysis.Rmd
Moved the exclusion block to lower in the script. Adjusted the disorder hierarchy section as a result of the change.

++MD: I had to cancel this chunk while running because it took too long - there are 473 variables in this dataset, so I don't think summarising all at once is useful!
```{r dat.coping summary, eval=FALSE, include=FALSE}
#summarytools::dfSummary(
#  dat.raw,
#  graph.col = F,
#  valid.col = F,
#  labels.col = F)
```

@@

```{r}
dat.raw %>%
  freq(Sample)
```

@@

Diagnoses analyses
Exclude individuals with NA for the MHD categories
```{r exclusion criteria for diagnoses analysis}
dat.mhd <- dat.raw %>%
  filter(
    !is.na(depressive_disorders) &
      !is.na(eating_disorders) &
      !is.na(psychotic_disorders) &
      !is.na(obsessive_compulsive_disorders) &
      !is.na(autism_spectrum_disorder) &
      !is.na(mhd.mania_hypomania_bipolar_or_manicdepression_numeric) &
      !is.na(mhd.posttraumatic_stress_disorder_ptsd_numeric)
  )

```{r disorders total count}
dat.raw %>%
  freq(disorders_total_count)
```

@@

## Hierarchy including psychosis
Diagnoses analyses
```{r Psychiatric disorders hierarchy including psychosis - dat.mhd}
dat.mhd <- dat.mhd %>%
  mutate(
    Disorder_hierarchical_psychosis_numeric =
      case_when(
        psychotic_disorders_numeric == 1 ~ 7,
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat.mhd <- dat.mhd %>%
  mutate(
    Disorder_hierarchical_psychosis =
      recode_factor(
        Disorder_hierarchical_psychosis_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders",
        `7` = "Psychotic disorders"
      )
  )

dat.mhd %>%
  freq(Disorder_hierarchical_psychosis)
```

Retro-baseline
```{r Psychiatric disorders hierarchy including psychosis - dat.retro}
dat.retro <- dat.retro %>%
  mutate(
    Disorder_hierarchical_psychosis_numeric =
      case_when(
        psychotic_disorders_numeric == 1 ~ 7,
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat.retro <- dat.retro %>%
  mutate(
    Disorder_hierarchical_psychosis =
      recode_factor(
        Disorder_hierarchical_psychosis_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders",
        `7` = "Psychotic disorders"
      )
  )

dat.retro %>%
  freq(Disorder_hierarchical_psychosis)
```

Prepan-baseline
++MD: For GLAD/EDGI we shouldn't have any controls, but some have crept in because they likely haven't ever received a diagnosis (despite qualifying on the CIDI or current measures of psychopathology). We should consider just excluding them.
```{r Psychiatric disorders hierarchy including psychosis - dat.prepan}
dat.prepan <- dat.prepan %>%
  mutate(
    Disorder_hierarchical_psychosis_numeric =
      case_when(
        psychotic_disorders_numeric == 1 ~ 7,
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat.prepan <- dat.prepan %>%
  mutate(
    Disorder_hierarchical_psychosis =
      recode_factor(
        Disorder_hierarchical_psychosis_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders",
        `7` = "Psychotic disorders"
      )
  )

dat.prepan %>%
  freq(Disorder_hierarchical_psychosis)
```

## Hierarchy NOT including psychosis
Diagnoses analyses
```{r Psychiatric disorders hierarchy without psychosis - dat.mhd}
dat.mhd <- dat.mhd %>%
  mutate(
    Disorder_hierarchical_numeric =
      case_when(
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat.mhd <- dat.mhd %>%
  mutate(
    Disorder_hierarchical =
      recode_factor(
        Disorder_hierarchical_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders"
      )
  )

dat.mhd %>%
  freq(Disorder_hierarchical) 
dat.mhd %>% 
  freq(Disorder_hierarchical_psychosis)
```

Retro-baseline
```{r Psychiatric disorders hierarchy without psychosis - dat.retro}
dat.retro <- dat.retro %>%
  mutate(
    Disorder_hierarchical_numeric =
      case_when(
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat.retro <- dat.retro %>%
  mutate(
    Disorder_hierarchical =
      recode_factor(
        Disorder_hierarchical_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders"
      )
  )

dat.retro %>%
  freq(Disorder_hierarchical)
dat.retro %>%
  freq(Disorder_hierarchical_psychosis)
```

Prepan-baseline
```{r Psychiatric disorders hierarchy without psychosis - dat.prepan}
dat.prepan <- dat.prepan %>%
  mutate(
    Disorder_hierarchical_numeric =
      case_when(
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat.prepan <- dat.prepan %>%
  mutate(
    Disorder_hierarchical =
      recode_factor(
        Disorder_hierarchical_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders"
      )
  )

dat.prepan %>%
  freq(Disorder_hierarchical)
dat.prepan %>%
  freq(Disorder_hierarchical_psychosis)
```

Investigate NAs 
Ask Leo?
++MD: you'll have NAs for individuals that don't have one of the listed disorders in the hierarchy, but who have another diagnosis that excludes them from being a control (e.g. personality disorder, ADHD, ASD). We can add an "Other" category to cover those other diagnoses, or exclude them from these analyses.
```{r NAs in diagnosis}
dat.mhd %>%
  filter(is.na(Disorder_hierarchical)) %>%
  freq(Sample)

dat.retro %>%
  filter(is.na(Disorder_hierarchical)) %>%
  freq(Sample)

dat.prepan %>%
  filter(is.na(Disorder_hierarchical)) %>%
  freq(Sample)
```

@@

@@

## Correlation matrix
++MD: Should we use the dat.mhd for this correlation matrix, as it has the least stringent inclusion criteria? Or just use dat.raw and include everyone? (alternatively, can create a correlation matrix for all 3 datasets)
```{r Correlation matrix of psychiatric disorders}
# Select items for correlation matrix
mhd.items <- c(
    "depressive_disorders",
    "anxiety_disorders",
    "eating_disorders",
    "obsessive_compulsive_disorders",
    "psychotic_disorders",
    "mhd.mania_hypomania_bipolar_or_manicdepression",
    "mhd.posttraumatic_stress_disorder_ptsd",
    "autism_spectrum_disorder",
    "mhd.attention_deficit_hyperactivity_disorder",
    "mhd.personality_disorder",
    "control"
)

>>>>

```{r dat.coping summary, eval=FALSE, include=FALSE}
#summarytools::dfSummary(
#  dat.raw,
#  graph.col = F,
#  valid.col = F,
#  labels.col = F)
```

@@

```{r sample frequencies}
dat.raw %>%
  freq(Sample)
```

@@

@@

## Hierarchy including psychosis
```{r Psychiatric disorders hierarchy including psychosis - dat.mhd}
dat.raw <- dat.raw %>%
  mutate(
    Disorder_hierarchical_psychosis_numeric =
      case_when(
        psychotic_disorders_numeric == 1 ~ 7,
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat.raw <- dat.raw %>%
  mutate(
    Disorder_hierarchical_psychosis =
      recode_factor(
        Disorder_hierarchical_psychosis_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders",
        `7` = "Psychotic disorders"
      )
  )

dat.raw %>%
  freq(Disorder_hierarchical_psychosis)
```

```{r Psychiatric disorders hierarchy without psychosis - dat.mhd}
dat.raw <- dat.raw %>%
  mutate(
    Disorder_hierarchical_numeric =
      case_when(
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat.raw <- dat.raw %>%
  mutate(
    Disorder_hierarchical =
      recode_factor(
        Disorder_hierarchical_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders"
      )
  )

dat.raw %>%
  freq(Disorder_hierarchical) 
dat.raw %>% 
  freq(Disorder_hierarchical_psychosis)
```

Investigate NAs 
NAs are coded for individuals that don't have one of the listed disorders in the hierarchy, but who have another diagnosis that excludes them from being a control (e.g. personality disorder, ADHD, ASD). These individuals will be excluded from the analyses.
```{r NAs in diagnosis}
dat.raw %>%
  filter(is.na(Disorder_hierarchical)) %>%
  freq(Sample)
```

Diagnoses analyses
Exclude individuals with NA for the MHD categories
```{r exclusion criteria for diagnoses analysis}
dat.mhd <- dat.raw %>%
  filter(
    !is.na(depressive_disorders) &
      !is.na(eating_disorders) &
      !is.na(psychotic_disorders) &
      !is.na(obsessive_compulsive_disorders) &
      !is.na(autism_spectrum_disorder) &
      !is.na(mhd.mania_hypomania_bipolar_or_manicdepression_numeric) &
      !is.na(mhd.posttraumatic_stress_disorder_ptsd_numeric) &
      !is.na(control)
  )

@@

```{r disorders total count}
dat.mhd %>%
  freq(disorders_total_count)
```

@@

## Correlation matrix
Using dat.mhd for the correlation matrix.
```{r Correlation matrix of psychiatric disorders}
# Select items for correlation matrix
mhd.items <- c(
    "depressive_disorders",
    "anxiety_disorders",
    "eating_disorders",
    "obsessive_compulsive_disorders",
    "psychotic_disorders",
    "mhd.mania_hypomania_bipolar_or_manicdepression",
    "mhd.posttraumatic_stress_disorder_ptsd",
    "autism_spectrum_disorder",
    "mhd.attention_deficit_hyperactivity_disorder",
    "mhd.personality_disorder",
    "control"
)
